---
title: "Medical Treatment Cost Prediction"
output: html_notebook
---

Importing Dataset
```{r}
data <- read.csv(file = "insurance.csv")
head(data)
```
Importing Required Libraries
```{r}
library(caTools) 
library(ggplot2)

```

Handling Categorial Variables
```{r}
data$gender <- ifelse(data$gender=="male",1,0)
head(data)
```


```{r}
data$smoker <- ifelse(data$smoker=="yes",1,0)
head(data)
```



```{r}
data$region<-ifelse(data$region=="southeast",1,0)
head(data)
```
Splitting Dataset
```{r}
set.seed(123)
split = sample.split(data$charges,SplitRatio = 1/2)
test_set = subset(data,split==TRUE)
training_set = subset(data,split==FALSE)
head(training_set)
head(test_set)
```
Checking for duplicate rows
```{r}
training_set[duplicated(training_set),]
```
checking for missing values
```{r}
colSums(is.na(training_set))
```
```{r}
colSums(is.na(test_set))
```
Some descriptive statistic
```{r}
summary(training_set)
```

Plotting Distribution of charges
```{r}
#png("charges_distribution.png")
charges_distribution <- ggplot(data=training_set,aes(x=charges))+geom_density(alpha=1)+ggtitle("Distribution Of Charges")
charges_distribution
#print(charges_distribution)
#dev.off()

```
Distribution is right skewed with a long tail to the right

```{r}
for (col in c('gender', 'region', 'children', 'smoker')) {
  plot <- ggplot(data = training_set,
                 aes_string(x = col, y = 'charges', group = col, fill = col)) + 
            geom_boxplot(show.legend = FALSE) + 
            ggtitle(glue::glue("Boxplot of Medical Charges per {col}"))
  print(plot)
}
```

```{r}

  charges_age_smoker <- ggplot(data = training_set, aes_string(x = 'age', y = 'charges', group = 'smoker', fill = 'smoker', col = 'smoker')) + 
    geom_jitter() + 
    geom_smooth(method = 'lm') +
    ggtitle(glue::glue("Charges vs Age"))  
charges_age_smoker

```
```{r}
 charges_bmi_smoker <- ggplot(data = training_set, aes_string(x = 'bmi', y = 'charges', group = 'smoker', fill = 'smoker', col = 'smoker')) + 
    geom_jitter() + 
    geom_smooth(method = 'lm') +
    ggtitle(glue::glue("Charges vs BMI"))  
  charges_bmi_smoker
```


```{r}
 charges_children_smoker <- ggplot(data = training_set, aes_string(x = 'children', y = 'charges', group = 'smoker', fill = 'smoker', col = 'smoker')) + 
    geom_jitter() + 
    geom_smooth(method = 'lm') +
    ggtitle(glue::glue("Charges vs Number of Children"))  
 charges_children_smoker
```


```{r}
regressor <- lm(formula = charges~.,data=training_set)
summary(regressor)
```

Excluding Insignificant Variables
```{r}
regressor <- lm(formula = charges~age+bmi+children+smoker,data=data)
summary(regressor)
```


```{r}
regressor <- lm(formula = charges~age+gender+bmi+children+smoker,data=data)
summary(regressor)
```


```{r}
regressor <- lm(formula = charges~age+region+bmi+children+smoker,data=data)
summary(regressor)
```


```{r}
plot(training_set$age,training_set$charges, xlab="Age", ylab="Charges", main="Age vs Charges",las=1, type="p")
```

Trying Polynomial Regression
```{r}
x <- data[-c(2,6,7)]
y <- data[7]
head(x)
```

```{r}
y <- data[7]
head(y)
```
Generating Polynomial Features
```{r}
pol <- poly(as.matrix(x[,1:4]) ,degree=2, raw=TRUE)
pol
```


```{r}
df <- data.frame(Age = pol[,1],
                  Age2=pol[,2],
                  Bmi=pol[,3],
                  Age_Bmi=pol[,4],
                  Bmi2=pol[,5],
                  Children=pol[,6],
                  Children_Age=pol[,7],
                  Children_bmi=pol[,8],
                  Children2= pol[,9],
                  Smoker=pol[,10],
                  Smoker_Age=pol[,11],
                  Smoker_Bmi=pol[,12],
                  Smoker_Children=pol[,13],
                  Smoker2=pol[,14])
```


```{r}
df_charges <- data.frame(y)
df$Charges <- df_charges[,1] 
df
```


```{r}
set.seed(123)
split = sample.split(df,SplitRatio = 1/5)
test = subset(df,split==TRUE)
train = subset(df,split==FALSE)
head(test)
head(train)
```


```{r}
pol_reg = lm(formula = Charges~., data = df)
summary(pol_reg)
```
```{r}
pol_reg1 = lm(formula = Charges~Age2+Bmi+Bmi2+Smoker+Smoker_Bmi, data = df)
summary(pol_reg1)
```


```{r}
pol_reg2 = lm(formula = Charges~Age2+Bmi+Bmi2+Smoker+Smoker_Bmi+Children, data = df)
summary(pol_reg2)

```


```{r}
pred <- predict(pol_reg2, newdata = test)
pred
```


```{r}
Charges_pred <- data.frame(pred)
df1 <- data.frame(Charges=test[,15])
df1$Predicted <- Charges_pred[,1]
df1
```


```{r}
p_age <- readline(prompt = "Enter Your Age: ")
p_age<- as.numeric(p_age)
p_bmi <- readline(prompt = "Enter Your BMI: ")
p_bmi<- as.numeric(p_bmi)
p_smoker <- readline(prompt = "Do You Smoke ? Enter 1 if Yes else 0: ")
p_smoker<- as.numeric(p_smoker)
p_children <- readline(prompt = "Enter Number of Children: ")
p_children<- as.numeric(p_children)
values <- data.frame(
              Age2 = p_age*p_age,
              Bmi = p_bmi,
              Bmi2 = p_bmi*p_bmi,
              Children=p_children,
              Smoker=p_smoker,
              Smoker_Bmi=p_smoker*p_bmi)
predicting <- predict(pol_reg2,newdata = values)
cat("Predicted Medical Treatment Cost: ",predicting)

```


```{r}
```

